### 2016-11-21 Paper_v1.tex
more changes in ssFBA MM

\subsection*{Estimation of kinetic model parameters.}
We estimated an ensemble of diverse parameter sets using a constrained Markov Chain Monte Carlo (MCMC) random walk strategy.
Starting from a single best fit parameter set estimated by inspection and literature,
we calculated the cost function, equal to the sum-squared-error between experimental data and model predictions:
\begin{equation}\label{eqn:cost-function}
    \texttt{cost}=\sum_{i=1}^{\mathcal{D}}\left[\frac{w_i}{\mathcal{Y}_{i}^2}\sum_{j=1}^{\mathcal{T}_i}\bigg(y_{ij}-x_{i}|_{t(j)}\bigg)^2 \right]
\end{equation}
where $\mathcal{D}$ denotes the number of datasets ($\mathcal{D}$~=~37),
$w_i$ denotes the weight of the $i^{th}$ dataset,
$\mathcal{T}_i$ denotes the number of timepoints in the $i^{th}$ dataset,
$t(j)$ denotes the $j^{th}$ timepoint, $y_{ij}$ denotes the measurement value of the $i^{th}$ dataset at the $j^{th}$ timepoint,
and $x_{i}|_{t(j)}$ denotes the simulated value of the metabolite corresponding to the $i^{th}$ dataset, interpolated to the $j^{th}$ timepoint.
Lastly, the cost calculation was scaled by the maximum experimental value in the $i^{th}$ dataset, $\mathcal{Y}_{i}=\max_{j}\left(y_{ij}\right)$.
We then perturbed each model parameter between an upper and lower bound that varied by parameter type:
\begin{equation}\label{eqn:parameter-perturbation}
    k_i^{new}=\min\left(\max\left(k_i \cdot \exp(a \cdot r_i),l_i\right),u_i\right)\qquad{i=1,2,\hdots,\mathcal{P}}
\end{equation}
where $\mathcal{P}$ denotes the number of parameters ($\mathcal{P}$~=~815), which includes 163 maximum reaction rates ($V^{max}$), 163 enzyme activity decay constants, 455 saturation constants ($K_{js}$), and 34 control parameters, $k_i^{new}$ denotes the new value of the $i^{th}$ parameter, $k_i$ denotes the current value of the $i^{th}$ parameter, $a$ denotes a distribution variance, $r_i$ denotes a random sample from the normal distribution, $l_i$ denotes the lower bound for that parameter type, and $u_i$ denotes the upper bound for that parameter type.
Maximum reaction rates were bounded between 0 and 500,000 mM/h \cite{Keseler:2013aa}.
Assuming a total enzyme concentration of 50 $\mu$M, this corresponds to catalytic rate bounds of 0 and 2778 s$^{-1}$.
These bounds resulted in a median catalytic rate of 0.016 s$^{-1}$ across the ensemble.
Enzyme activity decay constants were bounded between 0 and 1 h$^{-1}$, corresponding to half lives of 42 minutes and infinity; median = 25 min.
Saturation constants were bounded between 0.001 and 10 mM; median = 0.16 mM.
Control parameters (gains and orders) were left unbounded; gain median = 0.076, order median = 0.69.
For each newly generated parameter set, we re-solved the balance equations and calculated the cost function.
All sets with a lower cost (and some with higher cost) were accepted into the ensemble.
After generating greater that 10,000 sets, we selected N = 100 sets with minimal set to set correlation to avoid over-sampling any region of parameter space.
% The original 12,437-set ensemble had a mean Pearson correlation coefficient of 0.94 between pairs of sets, while the 100-set ensemble had a mean Pearson correlation coefficient of 0.83 between pairs of sets.

@@

\subsection*{Sequence specific calculation of carbon yield.}
We estimated the theoretical maximum carbon yield of CAT using sequence-specific flux balance analysis (ssFBA) \cite{Allen:2003aa}.
The CAT carbon yield ($Y_{C}^{CAT}$) was calculated as the ratio of carbon produced as CAT dived by the carbon consumed as reactants (glucose and amino acids):
\begin{equation}\label{eqn:yield-definition}
	Y_{C}^{CAT}=\frac{\Delta\texttt{CAT}\cdot C_{CAT}}{\displaystyle\sum_{i=1}^{\mathcal{R}}\max(\Delta m_{i},0)\cdot C_{m_i}}
\end{equation}
where $\Delta\texttt{CAT}$ denotes the abundance of CAT produced, $C_{CAT}$ denotes carbon number of CAT, $\mathcal{R}$ denotes the number of reactants,
$\Delta m_{i}$ denotes the amount of the $i^{th}$ reactant consumed (never allowed to be negative), and $C_{m_i}$ denotes the carbon number of the $i^{th}$ reactant.
Arginine or glutamate were not considered in the yield calculations, as no experimental measurements were available for these amino acids.
Yield of the best-fit parameter set and the experimental data were calculated by setting $\Delta\texttt{CAT}$ equal to the final minus the initial CAT concentration,
and setting $\Delta m_{i}$ equal to the initial minus the final reactant concentration.

The sequence specific flux balance analysis problem was formulated as:
\begin{equation}
 \begin{multlined}
	\qquad \qquad \qquad \max_{\boldsymbol{w}}{} \! \left( w_{obj} = \mathbf{\boldsymbol{\theta}}^T \boldsymbol{w} \right) \\
	\mathrm{Subject \; to:}
	 \; \; \mathbf{S}\mathbf{w}=\mathbf{0} \\
\alpha_i \leq w_i \leq \beta_i  \qquad i=1,2,\hdots,\mathcal{R}
 \end{multlined}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{w}$ denotes the unknown flux vector, $\boldsymbol{\theta}$ denotes the objective selection vector
and $\alpha_i$ and $\beta_i$ denote the lower and upper bounds on flux $w_{i}$, respectively.
The objective was to maximize the CAT translation rate.
The glucose uptake rate was bounded in the range [0,40 mM/h] according to experimental data;
while the amino acid uptake rates were bounded by [0,30 mM/h], but did not reach the maximum flux.
The transcription and translation rates were modeled using template reactions

The transcription rate ($w_{TX}$) was fixed in the ssFBA calculation as:
\begin{equation}
	w_{TX} = \left[R_{1}\left(\frac{K_{T7}}{1 + K_{T7}}\right)\left(\frac{v_{TX}}{l_{G}}\right)\right]\left(\frac{G}{K_{TX}+G}\right)
\end{equation}
where $R_{1}$ denotes the concentration of RNA polymerase, $v_{TX}$ denotes RNA polymerase elongation rate (nt/hr),
$l_{G}$ denotes the gene length in nucleotides, $G$ denotes the gene concentration and $K_{TX}$ denotes the plasmid saturation coefficient.
The gene and protein sequence for CAT was determined from literature.
The last term in the $w_{TX}$ expression describes T7 promoter activity, where $K_{T7}$ quantifies T7 RNA polymerase binding [REFHERE].
The translation rate ($w_{TL}$) was bounded by:
 \begin{equation}
	0\leq w_{TL} \leq \left[K_{P} R_{2}\left(\frac{v_{TL}}{l_{P}}\right)\right]\left(\frac{mRNA_{SS}}{K_{TL}+mRNA_{SS}}\right)
\end{equation}
where $K_{P}$ denotes the polysome amplification constant, $R_{2}$ denotes the ribosome concentration, $v_{TL}$ denotes the ribosome elongation rate (amino acids per hour),
$l_{P}$ denotes the number of amino acids in the protein of interest, and $mRNA_{SS}$ denotes the steady-state mRNA concentration:
\begin{equation}
	mRNA_{SS}\simeq\frac{w_{TX}}{\lambda}
\end{equation}where $\lambda$ denotes the mRNA degradation rate.

>>>>

\subsection*{Estimation of kinetic model parameters.}
We estimated an ensemble of diverse parameter sets using a constrained Markov Chain Monte Carlo (MCMC) random walk strategy.
Starting from a single best fit parameter set estimated by inspection and literature,
we calculated the cost function, equal to the sum-squared-error between experimental data and model predictions:
\begin{equation}\label{eqn:cost-function}
    \texttt{cost}=\sum_{i=1}^{\mathcal{D}}\left[\frac{w_i}{\mathcal{Y}_{i}^2}\sum_{j=1}^{\mathcal{T}_i}\bigg(y_{ij}-x_{i}|_{t(j)}\bigg)^2 \right]
\end{equation}
where $\mathcal{D}$ denotes the number of datasets ($\mathcal{D}$~=~37),
$w_i$ denotes the weight of the $i^{th}$ dataset,
$\mathcal{T}_i$ denotes the number of timepoints in the $i^{th}$ dataset,
$t(j)$ denotes the $j^{th}$ timepoint, $y_{ij}$ denotes the measurement value of the $i^{th}$ dataset at the $j^{th}$ timepoint,
and $x_{i}|_{t(j)}$ denotes the simulated value of the metabolite corresponding to the $i^{th}$ dataset, interpolated to the $j^{th}$ timepoint.
Lastly, the cost calculation was scaled by the maximum experimental value in the $i^{th}$ dataset, $\mathcal{Y}_{i}=\max_{j}\left(y_{ij}\right)$.
We then perturbed each model parameter between an upper and lower bound that varied by parameter type:
\begin{equation}\label{eqn:parameter-perturbation}
    k_i^{new}=\min\left(\max\left(k_i \cdot \exp(a \cdot r_i),l_i\right),u_i\right)\qquad{i=1,2,\hdots,\mathcal{P}}
\end{equation}
where $\mathcal{P}$ denotes the number of parameters ($\mathcal{P}$~=~815), which includes 163 maximum reaction rates ($V^{max}$), 163 enzyme activity decay constants, 455 saturation constants ($K_{js}$), and 34 control parameters, $k_i^{new}$ denotes the new value of the $i^{th}$ parameter, $k_i$ denotes the current value of the $i^{th}$ parameter, $a$ denotes a distribution variance, $r_i$ denotes a random sample from the normal distribution, $l_i$ denotes the lower bound for that parameter type, and $u_i$ denotes the upper bound for that parameter type.
Maximum reaction rates were bounded between 0 and 500,000 mM/h \cite{Keseler:2013aa}.
Assuming a total enzyme concentration of 5.0 $\mu$M, this corresponds to catalytic rate bounds of 0 and 27,780 s$^{-1}$.
These bounds resulted in a median catalytic rate of 0.16 s$^{-1}$ across the ensemble.
Enzyme activity decay constants were bounded between 0 and 1 h$^{-1}$, corresponding to half lives of 42 minutes and infinity; median = 25 h.
Saturation constants were bounded between 0.001 and 10 mM; median = 0.16 mM.
Control parameters (gains and orders) were left unbounded; gain median = 0.076, order median = 0.69.
For each newly generated parameter set, we re-solved the balance equations and calculated the cost function.
All sets with a lower cost (and some with higher cost) were accepted into the ensemble.
After generating greater that 10,000 sets, we selected N = 100 sets with minimal set to set correlation to avoid over-sampling any region of parameter space.
% The original 12,437-set ensemble had a mean Pearson correlation coefficient of 0.94 between pairs of sets, while the 100-set ensemble had a mean Pearson correlation coefficient of 0.83 between pairs of sets.

@@

\subsection*{Sequence specific calculation of carbon yield.}
We estimated the theoretical maximum CAT carbon yield using sequence-specific flux balance analysis (ssFBA) \cite{Allen:2003aa}.
The sequence specific flux balance analysis problem was formulated as:
\begin{equation}
 \begin{multlined}
	\qquad \qquad \qquad \max_{\boldsymbol{w}}{} \! \left( w_{TL} = \mathbf{\boldsymbol{\theta}}^T \boldsymbol{w} \right) \\
	\mathrm{Subject \; to:}
	 \; \; \mathbf{S}\mathbf{w}=\mathbf{0} \\
\alpha_i \leq w_i \leq \beta_i  \qquad i=1,2,\hdots,\mathcal{R}
 \end{multlined}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{w}$ denotes the unknown flux vector, $\boldsymbol{\theta}$ denotes the objective selection vector
and $\alpha_i$ and $\beta_i$ denote the lower and upper bounds on flux $w_{i}$, respectively.
The glucose uptake rate was bounded in the range [0,40 mM/h] according to experimental data;
while the amino acid uptake rates were bounded by [0,30 mM/h], but did not reach the maximum flux.
The transcription and translation rates were modeled using the template reactions: [FILL IN OR REFERENCE TEMPLATE TABLE].
The transcription rate ($w_{TX}$) was fixed in the ssFBA calculation as:
\begin{equation}
	w_{TX} = \left[R_{1}\left(\frac{v_{TX}}{l_{G}}\right)\left(\frac{K_{T7}}{1 + K_{T7}}\right)\right]\left(\frac{G}{K_{TX}+G}\right)
\end{equation}
The first term in the brackets denotes the maximum rate of transcription, where $R_{1}$ denotes the concentration of RNA polymerase, $v_{TX}$ denotes the RNA polymerase elongation rate (nt/hr), $l_{G}$ denotes the gene length in nucleotides, and the last term describes T7 promoter activity, where $K_{T7}$ quantifies T7 RNA polymerase binding [REF-HERE].
The second term denotes the gene saturation term, where $G$ denotes the gene concentration and $K_{TX}$ denotes a saturation coefficient.
The CAT gene and protein sequences were determined from literature.
The translation rate ($w_{TL}$) was bounded by:
 \begin{equation}
	0\leq w_{TL} \leq \left[K_{P} R_{2}\left(\frac{v_{TL}}{l_{P}}\right)\right]\left(\frac{mRNA_{SS}}{K_{TL}+mRNA_{SS}}\right)
\end{equation}
The first term in brackets denotes the maximum translation rate, where $K_{P}$ denotes the polysome amplification constant, $R_{2}$ denotes the ribosome concentration, $v_{TL}$ denotes the ribosome elongation rate (amino acids per hour), $l_{P}$ denotes the number of amino acids in the protein of interest, and $mRNA_{SS}$ denotes the steady-state mRNA concentration:
\begin{equation}
	mRNA_{SS}\simeq\frac{w_{TX}}{\lambda}
\end{equation}where $\lambda$ denotes the rate constant controlling the mRNA degradation rate.

The CAT carbon yield ($Y_{C}^{CAT}$) was calculated as the ratio of carbon produced as CAT dived by the carbon consumed as reactants (glucose and amino acids):
\begin{equation}\label{eqn:yield-definition}
	Y_{C}^{CAT}=\frac{\Delta\texttt{CAT}\cdot C_{CAT}}{\displaystyle\sum_{i=1}^{\mathcal{R}}\max(\Delta m_{i},0)\cdot C_{m_i}}
\end{equation}
where $\Delta\texttt{CAT}$ denotes the abundance of CAT produced, $C_{CAT}$ denotes carbon number of CAT, $\mathcal{R}$ denotes the number of reactants,
$\Delta m_{i}$ denotes the amount of the $i^{th}$ reactant consumed (never allowed to be negative), and $C_{m_i}$ denotes the carbon number of the $i^{th}$ reactant.
Arginine or glutamate were not considered in the yield calculations, as no experimental measurements were available for these amino acids.
Yield of the best-fit parameter set and the experimental data were calculated by setting $\Delta\texttt{CAT}$ equal to the final minus the initial CAT concentration,
and setting $\Delta m_{i}$ equal to the initial minus the final reactant concentration.
